---
title: "Project 1 - Hockey Vignette"
author: "Lucy Eckert"
date: "9/13/2020"
output:
  github_document:
    toc: true
    toc_depth: 4
---
#ST 558 Project #1 - Exploring the NHL API with the Boston Bruins

### First, a list of packages used in this exploration:
```{r}
library(tidyverse)
library(rmarkdown)
library(httr) 
library(jsonlite)
library(ggplot2)
library(knitr)
# rmarkdown::render("Project_1.Rmd", output_file="README.md")
```

### Let's dig into the API and take a look at accessing franchise data:
Return Parsed Franchise Data
```{r}
getFranchise <- function(name=NULL, ID=NULL){
  base_url <- "https://records.nhl.com/site/api"
  url <- paste0(base_url, "/", "franchise")
  dat <- as_tibble(fromJSON(content(GET(url),"text"),flatten = T)$data)
  if (!is.null(name)){
    dat <- dat %>% filter(teamCommonName %in% name)
  }
  if (!is.null(ID)){
  dat <- dat %>% filter(id %in% ID)
  }
  dat
}

FranchiseInfo <- getFranchise(name = "Bruins")
FranchiseInfo
```

Return Franchise Team Totals
```{r}
getFranTotal <- function(name=NULL, ID=NULL){
  base_url <- "https://records.nhl.com/site/api"
  url <- paste0(base_url, "/", "franchise-team-totals")
  dat <- as_tibble(fromJSON(content(GET(url),"text"),flatten = T)$data)
  if (!is.null(name)){
    dat <- dat %>% filter(teamName %in% name)
  }
  if (!is.null(ID)){
    dat <- dat %>% filter(teamId %in% ID)
  }
  dat
}

FranchiseTots <- getFranTotal(ID = 6)
FranchiseTots
```
 
Return Franchise Season Records
```{r}
getFranSeasonRecord <- function(name = NULL, ID=NULL){
  base_url <- "https://records.nhl.com/site/api"
  url <- paste0(base_url, "/", "franchise-season-records")
  if (is.null(ID)==F){
    url <- paste0(url, "?cayenneExp=franchiseId=", ID)
  }
  dat <- as_tibble(fromJSON(content(GET(url),"text"),flatten = T)$data)
  if (!is.null(name)){
    dat <- dat %>% filter(franchiseName %in% name)
  }
  if (!is.null(ID)){
    dat <- dat %>% filter(franchiseId %in% ID)
  }
  dat
}

Franch_SnsRcrds <- getFranSeasonRecord(name = "Boston Bruins")
Franch_SnsRcrds
```

Return Goalie Records
```{r}

getFranGoalieRecord <- function(name = NULL, ID=NULL){
  base_url <- "https://records.nhl.com/site/api"
  url <- paste0(base_url, "/", "franchise-goalie-records")
  if (is.null(ID)==F){
    url <- paste0(url, "?cayenneExp=franchiseId=", ID)
  }
  dat <- as_tibble(fromJSON(content(GET(url),"text"),flatten = T)$data)
   if (!is.null(name)){
    dat <- dat %>% filter(franchiseName %in% name)
  }
  if (!is.null(ID)){
    dat <- dat %>% filter(franchiseId %in% ID)
  }
  dat
}

Franch_GoalRcrds <- getFranGoalieRecord(name = "Boston Bruins")
Franch_GoalRcrds
```

Return Skater Records
```{r}
getFranSkaterRecord <- function(name = NULL, ID=NULL){
  base_url <- "https://records.nhl.com/site/api"
  url <- paste0(base_url, "/", "franchise-skater-records")
  if (is.null(ID)==F){
    url <- paste0(url, "?cayenneExp=franchiseId=", ID)
  }
  dat <- as_tibble(fromJSON(content(GET(url),"text"),flatten = T)$data)
  if (!is.null(name)){
    dat <- dat %>% filter(franchiseName %in% name)
  }
  if (!is.null(ID)){
    dat <- dat %>% filter(franchiseId %in% ID)
  }
  dat
}

Bruin_Sk8rRcrds <- getFranSkaterRecord(name = "Boston Bruins")
Bruin_Sk8rRcrds
```

Stats API - make a note that this only works for current teams
```{r}
getStatData <- function(name = NULL, expand=NULL, season=NULL, teamID=NULL, stats=NULL){
  url <- "https://statsapi.web.nhl.com/api/v1/teams"
  if (!is.null(expand)){
    url <- paste0(url, "?expand=", expand)
  }
  if (!is.null(season)){
    url <- paste0(url, "&season=", season)
  }
  if(!is.null(teamID)){
    teams <- paste(teamID, collapse = ",")
    url <- paste0(url, "?teamId=", teams)
  }
  if(!is.null(stats)){
    url <- paste0(url, "?stats=", stats)
  }
  dat <- as_tibble(fromJSON(content(GET(url),"text"),flatten = T)$teams)
  if (!is.null(name)){
    dat <- dat %>% filter(teamName %in% name)
  }
  if (!is.null(teamID)){
    dat <- dat %>% filter(franchiseId %in% teamID)
  }
  dat
}

F <- getStatData(teamID = 6)
F
```

wrapper function
```{r}
wrapper <- function(baseAPI="Record", EndPoint="Franchise", franID=NULL, name=NULL, 
                    expand=NULL, season=NULL, teamID=NULL, stats=NULL){
  
  if (toupper(baseAPI)=="RECORD"){
    if (grepl("SKATE", toupper(EndPoint))){
      dat <- getFranSkaterRecord(ID=franID)
    }
    else if (grepl("GOAL", toupper(EndPoint))){
      dat <- getFranGoalieRecord(ID=franID)
    }
    else if (grepl("SEASON", toupper(EndPoint))){
      dat <- getFranSeasonRecord(ID=franID)
    }
    else if (grepl("TOTAL", toupper(EndPoint))){
      dat <- getFranTotal()
    }
    else if (grepl("FRANCHISE", toupper(EndPoint))){
      dat <- getFranchise()
    }
  }
  else if (toupper(baseAPI)=="STATS"){
    dat <- getStatData(expand=expand, season=season, teamID=teamID, stats=stats)
  }
  else{
    stop("There is no function to call from different APIs")
  }
  dat
}
```

```{r}
plot_col <- ggplot(data = E, aes(x = division.name))
plot_col + geom_bar() + labs(x = "Teams per Division")
```

### Now that we have data, let's use it to answer some questions.  

#### Goals by Skater Position  

Not knowing much about hockey, I wondered which skating position usually scores the most goals. I used
my wrapper to pull all skater records, grouped by position, and then found the mean number of points
per season. This involved one extra step, which was creating a variable that calculated average
number of goals per season. 
```{r}
#what skater positions score most points
#first make new variable which is average points per season

positions <- wrapper(baseAPI="Record", EndPoint="skate", franID=NULL)
posit_andAvg <- mutate(positions, avgGls = (positions$goals/positions$seasons))
table1 <- posit_andAvg %>% group_by(positionCode) %>% summarise(Average = mean(avgGls))
kable(table1)
```

#### Get stats for current teams only  

First Ill the isolate only the current franchises. To do this, I'll pull the franchises
that are included in the Endpoint that lists upcoming games. Then, I'll join it with 
season record. We now have a massive dataset of current franchises full of lots of useful stats
```{r}
C <- getStatData(expand="team.schedule.next")
new <- inner_join(C, franch_seasonRecords, by = "franchiseId")
```


#### Are Pugnacious Bruins More Likely to Be High Scorers?  

```{r}
correlation <- cor(Bruin_Sk8rRcrds$penaltyMinutes, Bruin_Sk8rRcrds$points)
scatter <- ggplot(Bruin_Sk8rRcrds, aes(x = penaltyMinutes, y = points))
scatter + geom_point() + geom_smooth(method = lm, color = "yellow") + 
  geom_text(x = 1500, y = 1250, size = 5, label = paste0("Correlation = ", 
          round(correlation, 2))) + ylab("Total Career Points") + xlab("Total Career Penalty Minutes") +ggtitle("Correlation Between Career Points and Penalty Minutes")
```

#### Looking at Lifetime Stats by Skater Position  

```{r}
Bruin_Rcrds <- getFranSkaterRecord(name = "Boston Bruins")
selectBruins <- Bruin_Rcrds %>% select(penaltyMinutes, goals, seasons, assists, positionCode)
bruinstable <- function(position){
  data <- selectBruins %>% filter(positionCode == position) %>% select(-positionCode)
  kable(apply(data, 2, summary), caption = paste("Summary of Position", position), 
        digit = 1)
}
bruinstable("C")
bruinstable("D")
bruinstable("L")
bruinstable("R")
```

#### Does early suscess lead to career success?  

We can see that the majority of players are average caliber (average on the Bruins still being
an amazing athlete!), but that the players with "High" - over 41 rookie points, more likely to 
Rockstar players during their careers.

```{r, eval = FALSE}
#Dr. Post - these are rendering as a series of 0s and 1s, so I am adding eval = FALSE for now.
tbl1 <- table(Bruin_Sk8rRcrds$goals, Bruin_Sk8rRcrds$rookiePoints, Bruin_Sk8rRcrds$positionCode)
kable(tbl1[,,1], caption = "Centers")

tbl2 <- table(Bruin_Sk8rRcrds$goals, Bruin_Sk8rRcrds$rookiePoints, Bruin_Sk8rRcrds$positionCode)
kable(tbl2[,,2], caption = "Defense")

tbl3 <- table(Bruin_Sk8rRcrds$goals, Bruin_Sk8rRcrds$rookiePoints, Bruin_Sk8rRcrds$positionCode)
kable(tbl3[,,3], caption = "Left Wing")

tbl4 <- table(Bruin_Sk8rRcrds$goals, Bruin_Sk8rRcrds$rookiePoints, Bruin_Sk8rRcrds$positionCode)
kable(tbl4[,,4], caption = "Right Wing")
```

To illustrate the above summary tables graphically, I have constructed a bar graph:
```{r}
Bruin_Sk8rRcrds <- getFranSkaterRecord(name = "Boston Bruins")

Bruin_Sk8rRcrds$rookiePoints <- ifelse(Bruin_Sk8rRcrds$rookiePoints >= 41, "High",
                                       ifelse(Bruin_Sk8rRcrds$rookiePoints >= 21, "Elevated",
                                              ifelse(Bruin_Sk8rRcrds$rookiePoints >= 11, "Moderate", "Average")))

Bruin_Sk8rRcrds$goals <- ifelse(Bruin_Sk8rRcrds$goals >= 99, "Rockstar",
                                ifelse(Bruin_Sk8rRcrds$goals >= 31, "Star",
                                       ifelse(Bruin_Sk8rRcrds$goals >= 11, "Valuable", "Ok")))

ggplot(Bruin_Sk8rRcrds, aes(x = rookiePoints, na.rm = TRUE)) + geom_bar(aes(fill = goals), position = "stack") + 
  xlab("Boston Bruins Rookie Points") + scale_fill_discrete(name = "")
```


#### What positions usually get the most assists?  

We can see from the below boxplot that Center and Defense get more assists than Left Wing and
Right Wing. I was surprised by this as I'd assumed Defense didnt get a lot of assists.
```{r}
Bruin_Rcrds <- getFranSkaterRecord(name = "Boston Bruins")
ggplot(Bruin_Rcrds, aes(x = positionCode, y = mostAssistsOneSeason)) + geom_boxplot() + 
  geom_jitter(aes(color = positionCode)) + ggtitle("Boxplot for Career Goals") + scale_fill_discrete(name = "Positions", labels = c("Center", "Defense", "Left Wing", "Right Wing"))
```


